---
title: "P-model current run"
author: "Koen Hufkens and Beni Stocker"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{P-model current run}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[utf8]{inputenc}
---

This workflow runs the P-model with {rsofun} with the latest version and forcing as used in Stocker et al. (2020). Calibration and evaluation data is also as in Stocker et al. (2020). All data is from files created for Stocker et al. (2020).

```{r setup, include = FALSE}
# Install packages
# if ("rsofun" %in% installed.packages()) remove.packages("rsofun")
devtools::install_github(
  "geco-bern/rsofun",
  ref = "HEAD",
  upgrade = "never",
  force = TRUE
  )

install.packages("rpmodel")

# Load packages
library(rpmodel)
library(rsofun)
library(dplyr)
library(tidyr)
library(readr)
library(ggplot2)
library(lubridate)
library(knitr)
library(ggthemes)
library(here)

# Load functions
source("../R/eval_sofun.R")
source("../R/get_stats.R")
source("../R/analyse_modobs2.R")

# Set seed
set.seed(42)

# only select the sites from the New Phytologist paper
# load site list and filter on these (see below)
flue_sites <- readr::read_csv(here::here("data/flue_stocker18nphyt.csv")) %>%
  dplyr::filter( !is.na(cluster) ) %>%
  distinct(site) %>%
  pull(site)

# model forcing data
load(here("data/df_drivers_fluxnet2015.Rdata"))

df_drivers_fluxnet2015 <- df_drivers_fluxnet2015 %>%
  dplyr::select(sitename, forcing) %>%
  unnest(forcing) %>%
  dplyr::filter(!(month(date)==2 & mday(date)==29)) %>%

  ## model requires flux per seconds now
  mutate(prec = prec / (60*60*24), ppfd = ppfd / (60*60*24)) %>%

  ## assuming all precipitation in liquid form
  mutate(rainf = prec, snowf = 0) %>%

  ## required for new version, but not used because
  mutate(tmin = temp, tmax = temp) %>%

  group_by(sitename) %>%
  nest() %>%
  rename(forcing = data) %>%
  right_join(
    df_drivers_fluxnet2015 %>%
      dplyr::select(-forcing),
    by = "sitename"
  ) %>%
  ungroup() %>%
  rename(site_info = siteinfo, params_soil = df_soiltexture)

# site selection: good-quality sites from analysis of Stocker et al. (2018) New Phyt.
flue_sites <- readr::read_csv(here("data/flue_stocker18nphyt.csv")) %>%
  dplyr::filter( !is.na(cluster) ) %>%
  distinct(site) %>%
  pull(site)

# observational data as calibration target
# created from daily FLUXNE2015 data by:
# - using GPP_NT_VUT_REF
# - using QC>0.8
ddf_fluxnet_gpp <- readr::read_rds(here("data/ddf_fluxnet_gpp_v4.2.rds"))

# observational data for evaluation
load(here("data/obs_eval_fluxnet2015.Rdata"))

# site meta info
siteinfo_fluxnet2015 <- df_drivers_fluxnet2015 |>
  select(sitename, site_info) |>
  unnest(site_info)

# calibration sites
calibsites <- siteinfo_fluxnet2015 %>%
  dplyr::filter(!(sitename %in% c("DE-Akm", "IT-Ro1"))) %>%  # excluded because fapar data could not be downloaded (WEIRD)
  # dplyr::filter(!(sitename %in% c("AU-Wom"))) %>%  # excluded because no GPP data was found in FLUXNET file
  dplyr::filter(sitename != "FI-Sod") %>%  # excluded because some temperature data is missing
  dplyr::filter( c4 %in% c(FALSE, NA) & classid != "CRO" & classid != "WET" ) %>%
  dplyr::filter( sitename %in% flue_sites ) %>%
  pull(sitename)


# Calibrate model
siteinfo_fluxnet2015 <- df_drivers_fluxnet2015 |> 
  select(sitename, site_info) |> 
  unnest(site_info)

calibsites <- siteinfo_fluxnet2015 %>% 
  dplyr::filter(!(sitename %in% c("DE-Akm", "IT-Ro1"))) %>%  # excluded because fapar data could not be downloaded (WEIRD)
  # dplyr::filter(!(sitename %in% c("AU-Wom"))) %>%  # excluded because no GPP data was found in FLUXNET file
  dplyr::filter(sitename != "FI-Sod") %>%  # excluded because some temperature data is missing
  dplyr::filter( c4 %in% c(FALSE, NA) & classid != "CRO" & classid != "WET" ) %>%
  dplyr::filter( sitename %in% flue_sites ) %>%
  pull(sitename)

write_rds(calibsites, file = here("data/calibsites.rds"))

# GenSa calibration settings for v4.4
settings_calib <- list(
  method = "GenSA",
  metric = cost_rmse_pmodel,
  control = list(maxit = 100),
  par = list(
    kphio = list(lower = 0.03, upper = 0.15, init = 0.05),
    kphio_par_a = list(lower = -0.005, upper = 0, init = -0.0025),
    soilm_betao = list(lower = 0, upper = 1, init = 0.2),
    kc_jmax = list(lower = 0.2, upper = 0.8, init = 0.41)
  )
)

# fixed (not calibrated) model parameters
par_fixed = list(
  kphio_par_b        = 20,
  soilm_thetastar    = 0.6*240,
  beta_unitcostratio = 146.0,
  rd_to_vcmax        = 0.014,
  tau_acclim         = 30.0
  )

# # THIS IS NOT RUN, RUN BEFOREHAND AND SPECIFY PARAMETERS BY HAND BELOW
# # Calibrate the model.
# set.seed(1982)
# par_calib <- calib_sofun(
#   drivers = df_drivers_fluxnet2015 |>
#     dplyr::filter(sitename %in% calibsites) |>
#     dplyr::mutate(forcing = purrr::map(forcing, ~rename(., rain = rainf, snow = snowf))) |>
#     dplyr::mutate(forcing = purrr::map(forcing, ~mutate(., netrad = 0))),
#   obs = ddf_fluxnet_gpp,
#   settings = settings_calib,
#   par_fixed = par_fixed,
#   targets = "gpp"
#   # verbose = TRUE
# )
```

# Model parameters

Model parameters are determined for this version from calibration. The code for model calibration is not shown. See `vignettes/0_p-model_current_run.Rmd` for the code.

```{r}
params_modl <- list(
  kphio              = 0.04090209, # 0.09423773,  # par_calib$par["kphio"],
  kphio_par_a        = -0.001366283, # par_calib$par["kphio_par_a"],
  kphio_par_b        = par_fixed["kphio_par_b"],
  soilm_thetastar    = par_fixed["soilm_thetastar"],
  soilm_betao        = 0.683499654, # par_calib$par["soilm_betao"], # 0,  # 
  beta_unitcostratio = par_fixed["beta_unitcostratio"],
  rd_to_vcmax        = par_fixed["rd_to_vcmax"],
  tau_acclim         = par_fixed["tau_acclim"],
  kc_jmax            = 0.200000000 # par_calib$par["kc_jmax"] # 0.41  # 
)
```

# Model run

```{r warning = FALSE, message = FALSE}
output <- rsofun::runread_pmodel_f(
  df_drivers_fluxnet2015 |>
    dplyr::mutate(
      forcing = purrr::map(
        forcing,
        ~rename(., rain = rainf, snow = snowf))
      ) |>
    dplyr::mutate(
      forcing = purrr::map(
        forcing,
        ~mutate(., netrad = 0))
    ),
  par = params_modl
)
```

# Run evaluation

This runs the `eval_sofun()` routine included in the sofunCalVal package, and outputs daily, monthly and annual summary statistics comparing observed with simulated values. We'll retake this same routine when running on the latest release (i.e. the workflow is equivalent but for the release of the `rsofun` package)

```{r warning=FALSE, message=FALSE, error=FALSE, include = FALSE}
evalsites <- output %>%
  mutate(ntsteps = purrr::map_dbl(data, ~nrow(.))) %>%
  dplyr::filter(ntsteps > 0) %>%
  pull(sitename)

settings_eval <- list(
  benchmark = list( gpp = c("fluxnet") ),
  sitenames = evalsites,
  agg       = 8
)

##  Evaluate model ----
out_eval <- eval_sofun(
  output,
  settings_eval,
  obs_eval = obs_eval,
  overwrite = TRUE,
  light = FALSE
)
```

# Results

```{r warning=FALSE, message=FALSE, error=FALSE, echo = FALSE}
out_eval$gpp$fluxnet$metrics %>% 
  bind_rows(.id = "Level") %>% 
  kable()
```

## 8-daily, spatial and annual

```{r message = FALSE, warning = FALSE, echo = FALSE}
out_eval$gpp$fluxnet$plot$gg_modobs_xdaily
out_eval$gpp$fluxnet$plot$gg_modobs_spatial_annual
```

## Mean seasonal cycle

```{r message = FALSE, warning = FALSE, echo = FALSE}
out_eval$gpp$fluxnet$data$meandoydf_byclim %>% 
  dplyr::filter(climatezone %in% c(
    "Aw south", "BSk north", "Cfa north",
    "Cfb north", "Cfb south", "Csa north",
    "Csb north", "Dfb north", "Dfc north")
    ) %>%
  dplyr::filter(koeppen_code != "-") %>% 
  pivot_longer(
    c(obs_mean, mod_mean),
    names_to = "source",
    values_to = "gpp"
    ) %>% 
  ggplot() +
  geom_ribbon(
    aes(x = doy, ymin = obs_min, ymax = obs_max), 
    fill = "black", 
    alpha = 0.2
    ) +
  geom_line(aes(x = doy, y = gpp, color = source), size = 0.4) +
  labs(y = expression( paste("Simulated GPP (g C m"^-2, " d"^-1, ")" ) ), 
       x = "DOY") +
  facet_wrap( ~climatezone ) +
  theme_gray() +
  theme(legend.position = "bottom") +
  scale_color_manual(
    name="Setup: ",
    values=c("red", "black")
    )
```

## Drought response

```{r message = FALSE, warning = FALSE, echo = FALSE}
source("../R/align_events.R")
source("../R/eval_droughtresponse.R")

df_dday_agg <- eval_droughtresponse( 
  df = out_eval$gpp$fluxnet$data$ddf %>%
    rename(
      site = sitename
      ), 
  path_flue = "../data/flue_stocker18nphyt.csv",
  before=20,
  after=105,
  leng_threshold = 10, 
  nbins=10, 
  do_norm=TRUE
  )

df_dday_agg %>% 
  ggplot() +
  geom_hline(
    yintercept = 0,
    color = "black",
    linetype = "dotted"
    ) +
  geom_vline(
    xintercept = 0,
    color = "black",
    linetype = "dotted"
    ) +
  geom_line(
    aes(
      x = dday,
      y = median
      ),
    size = 0.9) +
  geom_ribbon(
    aes(
      x = dday,
      ymin = q33,
      ymax = q66
      ), 
    alpha = 0.3) +
  scale_color_manual(
    values = c(
      "BRC" = "black",
      "FULL" = "royalblue"
      ),
    name = "Setup"
    ) +
  scale_fill_manual(
    values = c(
      "BRC" = "black",
      "FULL" = "royalblue"
      ),
    name = "Setup") +
  ylim(-1.2, 2.2) + 
  xlim(-20, 105) +
  scale_x_continuous(
    expand = c(0,0)
    ) +
  scale_y_continuous(
    expand = c(0,0)
    ) +
  labs(
    x = "Days after drought onset",
    y = expression( paste( "Bias (g C m"^{-1}, " d"^{-1}, ")")) 
    ) +
  theme_classic()
```

## Consistency with rpmodel

Since rsofun takes time series forcing, overwrite forcing with constant values corresponding to the arguments provided to `rpmodel::rpmodel()`.

### Setup ORG

```{r  message = FALSE, warning = FALSE, include = FALSE}
library(rpmodel)
out_pmodel <- rpmodel( 
  tc             = 20,           # temperature, deg C
  vpd            = 1000,         # Pa,
  co2            = 400,          # ppm,
  fapar          = 1,            # fraction  ,
  ppfd           = 30,           # mol/m2/d,
  elv            = 0,            # m.a.s.l.,
  kphio          = 0.049977,     # quantum yield efficiency as calibrated
  beta           = 146,          # unit cost ratio a/b,
  c4             = FALSE,
  method_jmaxlim = "wang17",
  do_ftemp_kphio = FALSE,        # corresponding to setup ORG
  do_soilmstress = FALSE,        # corresponding to setup ORG
  verbose        = TRUE
  ) %>% 
  as_tibble()

## overwrite forcing with constant conditions (for one site)
drivers_rpmodel <- df_drivers_fluxnet2015 |> 
  plyr::mutate(forcing = purrr::map(forcing, ~rename(., rain = rainf, snow = snowf))) %>% 
  slice(1) %>% 
  mutate(forcing = purrr::map(forcing, ~mutate(., 
                                               temp = 20,
                                               vpd = 1000,
                                               ppfd = 30 / (60*60*24),
                                               patm = 101325,
                                               fapar = 1.0,
                                               co2 = 400,
                                               tmin = 20,
                                               tmax = 20))) %>% 
  mutate(params_siml = purrr::map(params_siml, ~mutate(.,
                                                       soilmstress = FALSE,
                                                       tempstress = FALSE)))

params_modl$kphio = 0.09423773 #pars$par[1]
params_modl$tau_acclim_tempstress = 1.0

out_rsofun <- rsofun::runread_pmodel_f(
  drivers_rpmodel,
  par = params_modl
  ) %>% 
  dplyr::select(data) %>%
  unnest(data) %>% 
  slice(1)
```

```{r  message = FALSE, warning = FALSE, echo = FALSE}
print("Are values equivalent for:")
paste("- ci:ca:", all.equal(out_pmodel$chi, out_rsofun$chi, tolerance = 1e-5))
paste("- GPP:", all.equal(out_pmodel$gpp, out_rsofun$gpp, tolerance = 1e-5))
paste("- Vcmax:", all.equal(out_pmodel$vcmax / (60*60*24), out_rsofun$vcmax, tolerance = 1e-5))
paste("- Vcmax25:", all.equal(out_pmodel$vcmax25 / (60*60*24), out_rsofun$vcmax25, tolerance = 1e-5))
paste("- Jmax:", all.equal(out_pmodel$jmax / (60*60*24), out_rsofun$jmax, tolerance = 1e-5))
paste("- Jmax25:", all.equal(out_pmodel$jmax25 / (60*60*24), out_rsofun$jmax25, tolerance = 1e-5))
```

### Setup BRC

```{r include = FALSE}
out_pmodel <- rpmodel( 
  tc             = 20,           # temperature, deg C
  vpd            = 1000,         # Pa,
  co2            = 400,          # ppm,
  fapar          = 1,            # fraction  ,
  ppfd           = 30,           # mol/m2/d,
  elv            = 0,            # m.a.s.l.,
  kphio          = 0.081785,     # quantum yield efficiency as calibrated for setup ORG by Stocker et al. 2020 GMD,
  beta           = 146,          # unit cost ratio a/b,
  c4             = FALSE,
  method_jmaxlim = "wang17",
  do_ftemp_kphio = TRUE,        # corresponding to setup ORG
  do_soilmstress = FALSE,        # corresponding to setup ORG
  verbose        = TRUE
  ) %>% 
  as_tibble()

## overwrite forcing with constant conditions (for one site)
drivers_rpmodel <- df_drivers_fluxnet2015 |> 
  plyr::mutate(forcing = purrr::map(forcing, ~rename(., rain = rainf, snow = snowf))) %>% 
  slice(1) %>% 
  mutate(forcing = purrr::map(forcing, ~mutate(., 
                                               temp = 20,
                                               vpd = 1000,
                                               ppfd = 30 / (60*60*24),
                                               patm = 101325,
                                               fapar = 1.0,
                                               co2 = 400,
                                               tmin = 20,
                                               tmax = 20))) %>% 
  mutate(params_siml = purrr::map(params_siml, ~mutate(.,
                                                       soilmstress = FALSE,
                                                       tempstress = FALSE)))

params_modl$kphio = 0.09423773 #pars$par[1]
params_modl$tau_acclim_tempstress = 1.0

out_rsofun <- rsofun::runread_pmodel_f(
  drivers_rpmodel,
  par = params_modl
  ) %>% 
  dplyr::select(data) %>% 
  unnest(data) %>% 
  slice(1)
```

```{r message = FALSE, warning = FALSE, echo = FALSE}
print("Are values equivalent for:")
paste("- ci:ca:", all.equal(out_pmodel$chi, out_rsofun$chi, tolerance = 1e-5))
paste("- GPP:", all.equal(out_pmodel$gpp, out_rsofun$gpp, tolerance = 1e-5))
paste("- Vcmax:", all.equal(out_pmodel$vcmax / (60*60*24), out_rsofun$vcmax, tolerance = 1e-5))
paste("- Vcmax25:", all.equal(out_pmodel$vcmax25 / (60*60*24), out_rsofun$vcmax25, tolerance = 1e-5))
paste("- Jmax:", all.equal(out_pmodel$jmax / (60*60*24), out_rsofun$jmax, tolerance = 1e-5))
paste("- Jmax25:", all.equal(out_pmodel$jmax25 / (60*60*24), out_rsofun$jmax25, tolerance = 1e-5))
```

## Appendix

```{r}
sessionInfo()
```
