

```{r setup, include = FALSE}
# Install packages
if ("rsofun" %in% installed.packages()) remove.packages("rsofun")
devtools::install_github(
  "geco-bern/rsofun@master",
  upgrade = "never",
  force = TRUE
  )
# Load packages
library(rsofun)
packageVersion("rsofun")
library(dplyr)
library(tidyr)
library(readr)
library(ggplot2)
library(lubridate)
library(knitr)
library(ggthemes)
library(here)

# Load functions
source("../R/eval_sofun.R")
source("../R/get_stats.R")
source("../R/analyse_modobs2.R")
source(here("R/align_events.R"))
source(here("R/eval_droughtresponse.R"))
source(here("R/heatscatter_dependencies.R"))


# Set seed
set.seed(42)


# model forcing data
driver <- read_rds("/data_2/FluxDataKit/v3.4/zenodo_upload/rsofun_driver_data_v3.4.rds") # currently only on workstation

# site meta info

fdk_site_info <- read_csv("/data_2/FluxDataKit/v3.4/zenodo_upload/fdk_site_info.csv")

fdk_filter <-  read_csv("/data_2/FluxDataKit/v3.4/zenodo_upload/fdk_site_fullyearsequence.csv")


fdk_site_info <- fdk_site_info[fdk_site_info$igbp_land_use != "CRO" &
                               fdk_site_info$igbp_land_use != "WET" 
                               # &
                               # fdk_site_info$sitename != "CA-SF1" &
                               # fdk_site_info$sitename != "CA-SF3" &
                               # fdk_site_info$sitename != "GF-Guy" &
                               # fdk_site_info$sitename != "ID-Pag" &
                               # fdk_site_info$sitename != "US-HWB" & 
                               # fdk_site_info$sitename != "GL-Dsk"
                               ,]

fdk_filter <- fdk_filter[fdk_filter$drop_gpp == "FALSE",]

driver <-driver[which(driver$sitename %in% fdk_site_info$sitename & 
                      driver$sitename %in% fdk_filter$sitename),]

fdk_site_info <-fdk_site_info[which(fdk_site_info$sitename %in% driver$sitename),]

fdk_filter <-fdk_filter[which(fdk_filter$sitename %in% driver$sitename &
                              fdk_filter$sitename %in% fdk_site_info$sitename),]

# select only year with good quality le

driver <- left_join(
  driver, fdk_filter |>
  select(sitename,year_start_gpp,year_end_gpp),
  by = "sitename") |>
  unnest(forcing) |>
  filter(between(as_date(date),
                 as_date(paste0(year_start_gpp,"-01-01")),
                 as_date(paste0(year_end_gpp,"-12-31"))
                 )
         ) |>
  group_by(sitename) |>
  nest(data = colnames(driver[1,]$forcing[[1]])) |>
  rename(forcing = data) |>
  select(-c(year_start_gpp,year_end_gpp))


# site selection: good-quality sites from analysis of Stocker et al. (2018) New Phyt.
df_flue <- readr::read_csv(here("data/flue_stocker18nphyt.csv"))

df_flue <- df_flue[df_flue$site %in% fdk_filter$sitename,]

df_flue <- left_join(df_flue |>
                       rename(sitename = site), fdk_filter |>
                       select(sitename,year_start_gpp,year_end_gpp),
                       by = "sitename") |>
           filter(between(as_date(date),
                 as_date(paste0(year_start_gpp,"-01-01")),
                 as_date(paste0(year_end_gpp,"-12-31"))
                 )) |>
           select(-c(year_start_gpp,year_end_gpp)) |>
           rename(site = sitename)


load(here("data/obs_eval_fluxnet2015.Rdata"), verbose = TRUE) # use as a blueprint
obs_eval$ddf <- NULL
obs_eval$xdf <- NULL
obs_eval$mdf <- NULL
obs_eval$adf <- NULL
obs_eval$breaks <- NULL


# daily data

ddf_tibble_preparation <- driver |>
  unnest(forcing) |>
  group_by(sitename) |>
  select(date,gpp) |>
  nest()

ddf_tibble_preparation <- ddf_tibble_preparation$data

obs_eval$ddf <- tibble(sitename = fdk_site_info$sitename,
                       lon = fdk_site_info$lon,
                       lat = fdk_site_info$lat,
                       elv = fdk_site_info$elv,
                       classid = fdk_site_info$igbp_land_use,
                       whc = fdk_site_info$whc,
                       koeppen_code = fdk_site_info$koeppen_code,
                       igbp_land_use = fdk_site_info$igbp_land_use,
                       plant_functioanl_type= NA,
                       c4 = FALSE,
                       data = ddf_tibble_preparation)

# weekly data

xdf_tibble_preparation <- driver |>
  unnest(forcing) |>
  mutate(week = lubridate::floor_date(date, "week")) |>
  select(sitename ,week,gpp) |>
  group_by(sitename, week) |>
  summarize(gpp = sum(gpp)) |>
  mutate(inbin = as.factor(week), date = week) |>
  select(date,inbin,gpp) |>
  nest()

xdf_tibble_preparation <- xdf_tibble_preparation$data

obs_eval$xdf <- tibble(sitename = fdk_site_info$sitename,
                       lon = fdk_site_info$lon,
                       lat = fdk_site_info$lat,
                       elv = fdk_site_info$elv,
                       classid = fdk_site_info$igbp_land_use,
                       whc = fdk_site_info$whc,
                       koeppen_code = fdk_site_info$koeppen_code,
                       igbp_land_use = fdk_site_info$igbp_land_use,
                       plant_functioanl_type= NA,
                       c4 = FALSE,
                       data = xdf_tibble_preparation)

# monthly data

mdf_tibble_preparation <- driver |>
  unnest(forcing) |>
  mutate(month = lubridate::floor_date(date, "month")) |>
  select(sitename ,month,gpp) |>
  group_by(sitename, month) |>
  summarize(gpp = sum(gpp)) |>
  mutate(date = month) |>
  select(date,gpp) |>
  nest()

mdf_tibble_preparation <- mdf_tibble_preparation$data

obs_eval$mdf <- tibble(sitename = fdk_site_info$sitename,
                       lon = fdk_site_info$lon,
                       lat = fdk_site_info$lat,
                       elv = fdk_site_info$elv,
                       classid = fdk_site_info$igbp_land_use,
                       whc = fdk_site_info$whc,
                       koeppen_code = fdk_site_info$koeppen_code,
                       igbp_land_use = fdk_site_info$igbp_land_use,
                       plant_functioanl_type= NA,
                       c4 = FALSE,
                       data = mdf_tibble_preparation)

# annual data

adf_tibble_preparation <- driver |>
  unnest(forcing) |>
  mutate(year = lubridate::floor_date(date, "year")) |>
  select(sitename ,year,gpp) |>
  group_by(sitename, year) |>
  summarize(gpp = sum(gpp)) |>
  mutate(date = year) |>
  select(date,gpp) |>
  nest()

adf_tibble_preparation <- adf_tibble_preparation$data

obs_eval$adf <- tibble(sitename = fdk_site_info$sitename,
                       lon = fdk_site_info$lon,
                       lat = fdk_site_info$lat,
                       elv = fdk_site_info$elv,
                       classid = fdk_site_info$igbp_land_use,
                       whc = fdk_site_info$whc,
                       koeppen_code = fdk_site_info$koeppen_code,
                       igbp_land_use = fdk_site_info$igbp_land_use,
                       plant_functioanl_type= NA,
                       c4 = FALSE,
                       data = adf_tibble_preparation)


breaks_preparation <- driver |>
  unnest(forcing) |>
  select(date) |>
  arrange(date) |>
  distinct(date) |>
  mutate(breaks = lubridate::floor_date(date, "week"))


obs_eval$breaks <- unique((breaks_preparation$breaks))[-1]


# # THIS IS NOT RUN, RUN BEFOREHAND AND SPECIFY PARAMETERS BY HAND BELOW
# set.seed(1982)
# pars <- calib_sofun(
#     drivers  = df_drivers_fluxnet2015 |> 
#       dplyr::filter(sitename %in% calibsites) |> 
#       dplyr::mutate(forcing = purrr::map(forcing, ~rename(., rain = rainf, snow = snowf))),
#   obs      = ddf_fluxnet_gpp,
#   settings = settings_calib
#   )

```

# Model parameters

Model parameters are determined for this version from calibration. The code for model calibration is not shown. See `vignettes/1_p-model_reference_run.Rmd` for the code.

```{r}
params_modl <- list(
    kphio              = 0.04998,    # setup ORG in Stocker et al. 2020 GMD
    kphio_par_a        = 0.0,        # set to zero to disable temperature-dependence of kphio
    kphio_par_b        = 1.0,
    soilm_thetastar    = 0.6 * 240,  # to recover old setup with soil moisture stress
    soilm_betao        = 0.0,
    beta_unitcostratio = 146.0,
    rd_to_vcmax        = 0.014,      # value from Atkin et al. 2015 for C3 herbaceous
    tau_acclim         = 30.0,
    kc_jmax            = 0.41
  )
  
output <- NULL

for(i in 1:nrow(driver)){
  
  tmp <- rsofun::runread_pmodel_f(
  driver[i,],
  par = params_modl
  )
  output <- rbind(output, tmp)
  
}

# TODO: some sites do not work here:
output |>
  mutate(len = purrr::map_int(data, ~nrow(.))) |>
  filter(len == 1) |>
  select(-len)

output <- output |>
  mutate(len = purrr::map_int(data, ~nrow(.))) |>
  filter(len != 1) |>
  select(-len)
```

# Model run


# Run evaluation

This runs the `eval_sofun()` routine included in the sofunCalVal package, and outputs daily, monthly and annual summary statistics comparing observed with simulated values. We'll retake this same routine when running on the latest release (i.e. the workflow is equivalent but for the release of the `rsofun` package)

```{r warning=FALSE, message=FALSE, error=FALSE, include = FALSE}
evalsites <- output %>% 
  mutate(ntsteps = purrr::map_dbl(data, ~nrow(.))) %>% 
  dplyr::filter(ntsteps > 0) %>% 
  pull(sitename)

# write_rds(evalsites, file = here("data/evalsites.rds"))

settings_eval <- list(
  benchmark = list( gpp = c("fluxnet") ),
  sitenames = evalsites,
  agg       = 8
  )

out_eval <- eval_sofun( 
  output, 
  settings_eval, 
  obs_eval = obs_eval, 
  overwrite = TRUE, 
  light = FALSE 
  )
```

# Results

```{r warning=FALSE, message=FALSE, error=FALSE, echo = FALSE}
out_eval$gpp$fluxnet$metrics %>% 
  bind_rows(.id = "Level") %>% 
  kable()
```

## 8-daily, spatial and annual

```{r message = FALSE, warning = FALSE, echo = FALSE}
out_eval$gpp$fluxnet$plot$gg_modobs_xdaily
out_eval$gpp$fluxnet$plot$gg_modobs_spatial_annual
```

## Mean seasonal cycle

```{r message = FALSE, warning = FALSE, echo = FALSE}
out_eval$gpp$fluxnet$data$meandoydf_byclim %>% 
  dplyr::filter(climatezone %in% c(
    "Aw south", "BSk north", "Cfa north",
    "Cfb north", "Cfb south", "Csa north",
    "Csb north", "Dfb north", "Dfc north")
    ) %>%
  dplyr::filter(koeppen_code != "-") %>% 
  pivot_longer(
    c(obs_mean, mod_mean),
    names_to = "source",
    values_to = "gpp"
    ) %>% 
  ggplot() +
  geom_ribbon(
    aes(x = doy, ymin = obs_min, ymax = obs_max), 
    fill = "black", 
    alpha = 0.2
    ) +
  geom_line(aes(x = doy, y = gpp, color = source), size = 0.4) +
  labs(y = expression( paste("Simulated GPP (g C m"^-2, " d"^-1, ")" ) ), 
       x = "DOY") +
  facet_wrap( ~climatezone ) +
  theme_gray() +
  theme(legend.position = "bottom") +
  scale_color_manual(
    name = "Setup: ",
    values = c("red", "black")
    )
```

## Drought response

```{r message = FALSE, warning = FALSE, echo = FALSE}
df_dday_agg <- eval_droughtresponse( 
  df = out_eval$gpp$fluxnet$data$ddf %>%
    rename(
      site = sitename
      ), 
  df_flue = readr::read_csv(here::here("data/flue_stocker18nphyt.csv")),
  before = 20,
  after = 105,
  leng_threshold = 10, 
  nbins = 10, 
  do_norm = TRUE
  )

df_dday_agg %>% 
  ggplot() +
  geom_hline(
    yintercept = 0,
    color = "black",
    linetype = "dotted"
    ) +
  geom_vline(
    xintercept = 0,
    color = "black",
    linetype = "dotted"
    ) +
  geom_line(
    aes(
      x = dday,
      y = median
      ),
    size = 0.9) +
  geom_ribbon(
    aes(
      x = dday,
      ymin = q33,
      ymax = q66
      ), 
    alpha = 0.3) +
  scale_color_manual(
    values = c(
      "BRC" = "black",
      "FULL" = "royalblue"
      ),
    name = "Setup"
    ) +
  scale_fill_manual(
    values = c(
      "BRC" = "black",
      "FULL" = "royalblue"
      ),
    name = "Setup") +
  ylim(-1.2, 2.2) + 
  xlim(-20, 105) +
  scale_x_continuous(
    expand = c(0,0)
    ) +
  scale_y_continuous(
    expand = c(0,0)
    ) +
  labs(
    x = "Days after drought onset",
    y = expression( paste( "Bias (g C m"^{-1}, " d"^{-1}, ")")) 
    ) +
  theme_classic() +
  theme(
    panel.grid.major.y = element_line(),
    panel.grid.major.x = element_line()
  )
```

## Appendix

```{r}
sessionInfo()
```
