

```{r setup, include = FALSE}
# Install packages
# if ("rsofun" %in% installed.packages()) remove.packages("rsofun")
# devtools::install_github(
#   "geco-bern/rsofun@master",
#   upgrade = "never",
#   force = TRUE
#   )

# Load packages
library(rsofun)
packageVersion("rsofun")
library(dplyr)
library(tidyr)
library(readr)
library(ggplot2)
library(lubridate)
library(knitr)
library(ggthemes)
library(here)

# Load functions
source("../R/eval_sofun.R")
source("../R/get_stats.R")
source("../R/analyse_modobs2.R")
source(here("R/align_events.R"))
source(here("R/eval_droughtresponse.R"))
source(here("R/heatscatter_dependencies.R"))
source("../R/create_obs_eval.R")

# Set seed
set.seed(42)

# model forcing data
driver <- read_rds("~/data_2/FluxDataKit/v3.4/zenodo_upload/rsofun_driver_data_v3.4.rds") # currently only on workstation

# site meta info
fdk_site_info <- read_csv("~/data_2/FluxDataKit/v3.4/zenodo_upload/fdk_site_info.csv")

# data quality filter info
fdk_filter <-  read_csv("~/data_2/FluxDataKit/v3.4/zenodo_upload/fdk_site_fullyearsequence.csv")

# exclude croplands and wetlands from calibration/evaluation
fdk_site_info <- fdk_site_info[fdk_site_info$igbp_land_use != "CRO" &
                               fdk_site_info$igbp_land_use != "WET" 
                               # &
                               # fdk_site_info$sitename != "CA-SF1" &
                               # fdk_site_info$sitename != "CA-SF3" &
                               # fdk_site_info$sitename != "GF-Guy" &
                               # fdk_site_info$sitename != "ID-Pag" &
                               # fdk_site_info$sitename != "US-HWB" & 
                               # fdk_site_info$sitename != "GL-Dsk"
                               ,]

# apply good year sequence data quality filter for GPP
fdk_filter <- fdk_filter[fdk_filter$drop_gpp == "FALSE",]

driver <- driver[which(driver$sitename %in% fdk_site_info$sitename &
                       driver$sitename %in% fdk_filter$sitename),]

fdk_site_info <- fdk_site_info[which(fdk_site_info$sitename %in% driver$sitename),]

fdk_filter <- fdk_filter[which(fdk_filter$sitename %in% driver$sitename &
                               fdk_filter$sitename %in% fdk_site_info$sitename),]

driver_forcing <- driver |> 
  select(sitename, forcing) |> 
  unnest(cols = c(forcing))

driver <- driver_forcing |>
  left_join(
    fdk_filter |> 
      select(
        sitename, 
        year_start = year_start_gpp, 
        year_end = year_end_gpp),
    by = join_by(sitename)
  ) |> 
  mutate(year = year(date)) |> 
  filter(year >= year_start & year <= year_end) |> 
  select(-year_start, -year_end, -year) |> 
  group_by(sitename) |> 
  nest() |> 
  left_join(
    driver |> 
      select(
        sitename, 
        site_info,
        params_siml
      ),
    by = join_by(sitename)
  ) |> 
  rename(forcing = data) |> 
  select(sitename, params_siml, site_info, forcing) |> 
  ungroup()

# site selection: good-quality sites from analysis of Stocker et al. (2018) New Phyt.
df_flue <- readr::read_csv(here("data/flue_stocker18nphyt.csv"))

df_flue <- df_flue[df_flue$site %in% fdk_filter$sitename,]

df_flue <- left_join(df_flue |>
                       rename(sitename = site), fdk_filter |>
                       select(sitename,year_start_gpp,year_end_gpp),
                       by = "sitename") |>
           filter(between(as_date(date),
                 as_date(paste0(year_start_gpp,"-01-01")),
                 as_date(paste0(year_end_gpp,"-12-31"))
                 )) |>
           select(-c(year_start_gpp,year_end_gpp)) |>
           rename(site = sitename)

# read fixed site split from file here, file created by a separate script 

# calibrate model using code in analysis/calibration_gensa_new_data_master_50.R
# this writes calibrated parameters to file data/calibration_new_data_50.rds
# that file also contains the list of sites used for model training as list element $sitename

par_calib <-  read_rds(here("data","calibration_new_data_50.rds"))

# construct list of observations used for evaluation
# xxx creates an error
obs_eval <- create_obs_eval(
  driver |> 
    filter(!(sitename %in% par_calib$sitename)),
  fdk_site_info |> 
    filter(!(sitename %in% par_calib$sitename))
  )
```

# Model parameters

Model parameters are determined for this version from calibration. The code for model calibration is not shown. See `analysis/analysis/calibration_gensa_new_data_master_50.R` for the code.

```{r}
params_modl <- list(
  kphio              = par_calib$par[["kphio"]],    
  kphio_par_a        = par_calib$par[["kphio_par_a"]],       
  kphio_par_b        = par_calib$par[["kphio_par_b"]],
  soilm_thetastar    = par_calib$par[["soilm_thetastar"]],  
  soilm_betao        = 0.0,
  beta_unitcostratio = 146.0,
  rd_to_vcmax        = 0.014,      # value from Atkin et al. 2015 for C3 herbaceous
  tau_acclim         = 30.0,
  kc_jmax            = 0.41
  )

output <- runread_pmodel_f(
  driver |> 
    filter(!(sitename %in% par_calib$sitename)),
  par = params_modl
  )

# TODO: some sites do not work here:
output |>
  mutate(len = purrr::map_int(data, ~nrow(.))) |>
  filter(len == 1) |>
  select(-len)

output <- output |>
  mutate(len = purrr::map_int(data, ~nrow(.))) |>
  filter(len != 1) |>
  select(-len)
```

# Model run

# Run evaluation

This runs the `eval_sofun()` routine included in the sofunCalVal package, and outputs daily, monthly and annual summary statistics comparing observed with simulated values. We'll retake this same routine when running on the latest release (i.e. the workflow is equivalent but for the release of the `rsofun` package)

```{r warning=FALSE, message=FALSE, error=FALSE, include = FALSE}
evalsites <- output$sitename[!(output$sitename %in% par_calib$sitename )]

settings_eval <- list(
  benchmark = list( gpp = c("fluxnet") ),
  sitenames = evalsites,
  agg       = 8
  )

out_eval <- eval_sofun( 
  output, 
  settings_eval, 
  obs_eval = obs_eval, 
  overwrite = TRUE, 
  light = FALSE 
  )
```

# Results

```{r warning=FALSE, message=FALSE, error=FALSE, echo = FALSE}
out_eval$gpp$fluxnet$metrics %>% 
  bind_rows(.id = "Level") %>% 
  kable()
```

## 8-daily, spatial and annual

<!-- xxx plot means across weeks, months, years, ... -->

```{r message = FALSE, warning = FALSE, echo = FALSE}
out_eval$gpp$fluxnet$plot$gg_modobs_xdaily
out_eval$gpp$fluxnet$plot$gg_modobs_spatial_annual
```

## Mean seasonal cycle

```{r message = FALSE, warning = FALSE, echo = FALSE}
out_eval$gpp$fluxnet$data$meandoydf_byclim %>% 
  dplyr::filter(climatezone %in% c(
    "Aw south", "BSk north", "Cfa north",
    "Cfb north", "Cfb south", "Csa north",
    "Csb north", "Dfb north", "Dfc north")
    ) %>%
  dplyr::filter(koeppen_code != "-") %>% 
  pivot_longer(
    c(obs_mean, mod_mean),
    names_to = "source",
    values_to = "gpp"
    ) %>% 
  ggplot() +
  geom_ribbon(
    aes(x = doy, ymin = obs_min, ymax = obs_max), 
    fill = "black", 
    alpha = 0.2
    ) +
  geom_line(aes(x = doy, y = gpp, color = source), size = 0.4) +
  labs(y = expression( paste("Simulated GPP (g C m"^-2, " d"^-1, ")" ) ), 
       x = "DOY") +
  facet_wrap( ~climatezone ) +
  theme_gray() +
  theme(legend.position = "bottom") +
  scale_color_manual(
    name = "Setup: ",
    values = c("red", "black")
    )
```

## Drought response

```{r message = FALSE, warning = FALSE, echo = FALSE}
df_dday_agg <- eval_droughtresponse( 
  df = out_eval$gpp$fluxnet$data$ddf %>%
    rename(
      site = sitename
      ), 
  df_flue = readr::read_csv(here::here("data/flue_stocker18nphyt.csv")),
  before = 20,
  after = 105,
  leng_threshold = 10, 
  nbins = 10, 
  do_norm = TRUE
  )

df_dday_agg %>% 
  ggplot() +
  geom_hline(
    yintercept = 0,
    color = "black",
    linetype = "dotted"
    ) +
  geom_vline(
    xintercept = 0,
    color = "black",
    linetype = "dotted"
    ) +
  geom_line(
    aes(
      x = dday,
      y = median
      ),
    size = 0.9) +
  geom_ribbon(
    aes(
      x = dday,
      ymin = q33,
      ymax = q66
      ), 
    alpha = 0.3) +
  scale_color_manual(
    values = c(
      "BRC" = "black",
      "FULL" = "royalblue"
      ),
    name = "Setup"
    ) +
  scale_fill_manual(
    values = c(
      "BRC" = "black",
      "FULL" = "royalblue"
      ),
    name = "Setup") +
  ylim(-1.2, 2.2) + 
  xlim(-20, 105) +
  scale_x_continuous(
    expand = c(0,0)
    ) +
  scale_y_continuous(
    expand = c(0,0)
    ) +
  labs(
    x = "Days after drought onset",
    y = expression( paste( "Bias (g C m"^{-1}, " d"^{-1}, ")")) 
    ) +
  theme_classic() +
  theme(
    panel.grid.major.y = element_line(),
    panel.grid.major.x = element_line()
  )
```

## Appendix

```{r}
sessionInfo()
```
