
```{r setup, include = FALSE}
# Install packages
 if ("rsofun" %in% installed.packages()) remove.packages("rsofun")
 devtools::install_github(
   "geco-bern/rsofun@phydro",
   upgrade = "never",
   force = TRUE
   )

# Load packages
library(rsofun)
packageVersion("rsofun")
library(dplyr)
library(tidyr)
library(readr)
library(ggplot2)
library(lubridate)
library(knitr)
library(ggthemes)
library(here)
library(ncdf4)

# Load functions
source("../R/eval_sofun.R")
source("../R/get_stats.R")
source("../R/analyse_modobs2.R")
source(here("R/align_events.R"))
source(here("R/eval_droughtresponse.R"))
source(here("R/heatscatter_dependencies.R"))
source(here("R/create_obs_eval.R"))


# Set seed
set.seed(42)

# model forcing data
driver <- read_rds("/data_2/FluxDataKit/v3.4/zenodo_upload/rsofun_driver_data_v3.4.rds") # currently only on workstation

# site meta info

fdk_site_info <- read_csv("/data_2/FluxDataKit/v3.4/zenodo_upload/fdk_site_info.csv")

fdk_filter <-  read_csv("/data_2/FluxDataKit/v3.4/zenodo_upload/fdk_site_fullyearsequence.csv")


fdk_site_info <- fdk_site_info[fdk_site_info$igbp_land_use != "CRO" &
                               fdk_site_info$igbp_land_use != "WET" 
                               # &
                               # fdk_site_info$sitename != "CA-SF1" &
                               # fdk_site_info$sitename != "CA-SF3" &
                               # fdk_site_info$sitename != "GF-Guy" &
                               # fdk_site_info$sitename != "ID-Pag" &
                               # fdk_site_info$sitename != "US-HWB" & 
                               # fdk_site_info$sitename != "GL-Dsk"
                               ,]

fdk_filter <- fdk_filter[fdk_filter$drop_gpp == "FALSE",]

driver <-driver[which(driver$sitename %in% fdk_site_info$sitename & 
                      driver$sitename %in% fdk_filter$sitename),]

fdk_site_info <-fdk_site_info[which(fdk_site_info$sitename %in% driver$sitename),]

fdk_filter <-fdk_filter[which(fdk_filter$sitename %in% driver$sitename &
                              fdk_filter$sitename %in% fdk_site_info$sitename),]

# select only year with good quality le

# driver <- left_join(
#   driver, fdk_filter |>
#   select(sitename,year_start_gpp,year_end_gpp),
#   by = "sitename") |>
#   unnest(forcing) |>
#   filter(between(as_date(date),
#                  as_date(paste0(year_start_gpp,"-01-01")),
#                  as_date(paste0(year_end_gpp,"-12-31"))
#                  )
#          ) |>
#   group_by(sitename) |>
#   nest(data = colnames(driver[1,]$forcing[[1]])) |>
#   rename(forcing = data) |>
#   select(-c(year_start_gpp,year_end_gpp))


# site selection: good-quality sites from analysis of Stocker et al. (2018) New Phyt.
df_flue <- readr::read_csv(here("data/flue_stocker18nphyt.csv"))

df_flue <- df_flue[df_flue$site %in% fdk_filter$sitename,]

df_flue <- left_join(df_flue |>
                       rename(sitename = site), fdk_filter |>
                       select(sitename,year_start_gpp,year_end_gpp),
                       by = "sitename") |>
           filter(between(as_date(date),
                 as_date(paste0(year_start_gpp,"-01-01")),
                 as_date(paste0(year_end_gpp,"-12-31"))
                 )) |>
           select(-c(year_start_gpp,year_end_gpp)) |>
           rename(site = sitename)


nc_file <- nc_open("~/data_scratch/sofuncalvalGPP/sofuncalvalGPP/data/whc_2m.nc")

whc = ncvar_get(nc_file, "whc_2m")
lons = ncvar_get(nc_file, "lon")
lats = ncvar_get(nc_file, "lat")

geo <- driver |>
  unnest(site_info) |>
  select(lon  , lat)

geo$sitename <- driver$sitename

n <- 1 # parameter to select size of slice to average

old_whc <- lapply(geo$sitename, function(x){
  tmp <- geo[geo$sitename == x,]
  lonid <- which(lons > tmp$lon)[1]
  latid <- which(lats > tmp$lat)[1]
  whc_grid <- whc[(lonid-n):(lonid+n), (latid-n):(latid+n)]
  whc_site <- mean(as.numeric(whc_grid, na.rm=T))
  return(whc_site)
})

old_whc = unlist(old_whc)

for(i in 1:dim(driver)[1]){
  driver$site_info[i][[1]][4] <- old_whc[i]
}



obs_eval <- create_obs_eval(driver,fdk_site_info)
```

# Model parameters

Model parameters are determined for this version from calibration. The code for model calibration is not shown. 

# Model run

```{r warning = FALSE, message = FALSE}

par_calib <-  read_rds(here("data","calibration_new_data_phydro_50.rds"))

params_modl <- list(
    kphio              = par_calib$par[["kphio"]],   
    kphio_par_a        = par_calib$par[["kphio_par_a"]],       
    kphio_par_b        = par_calib$par[["kphio_par_b"]],
    rd_to_vcmax        = 0.014,  
    soilm_thetastar    = par_calib$par[["soilm_thetastar"]], 
    beta_unitcostratio = 146,
    tau_acclim         = 30,
    kc_jmax            = 0.54131889
  )

driver <- driver |> unnest(params_siml) |>
  mutate(use_gs = FALSE,
         use_phydro = FALSE,
         use_pml= FALSE) |>
  group_by(sitename) |>
  nest(params_siml = c(spinup, spinupyears, recycle, outdt, ltre,  ltne,  ltrd,  ltnd,  lgr3,  lgn3,  lgr4 ,
                       use_gs, use_phydro, use_pml))
  
 output <- rsofun::runread_pmodel_f(
  driver,
  par = params_modl
 )

# remove sites where it didn't work
output <- output |>
  mutate(len = purrr::map_int(data, ~nrow(.))) |>
  filter(len != 1) |>
  select(-len)
```

# Run evaluation

This runs the `eval_sofun()` routine included in the sofunCalVal package, and outputs daily, monthly and annual summary statistics comparing observed with simulated values. We'll retake this same routine when running on the latest release (i.e. the workflow is equivalent but for the release of the `rsofun` package)

```{r warning=FALSE, message=FALSE, error=FALSE, include = FALSE}
evalsites <- output$sitename[!(output$sitename %in% par_calib$sitename )]

settings_eval <- list(
  benchmark = list( gpp = c("fluxnet") ),
  sitenames = evalsites,
  agg       = 8
  )

out_eval <- eval_sofun( 
  output, 
  settings_eval, 
  obs_eval = obs_eval, 
  overwrite = TRUE, 
  light = FALSE 
  )
```

# Results

```{r warning=FALSE, message=FALSE, error=FALSE, echo = FALSE}
out_eval$gpp$fluxnet$metrics %>% 
  bind_rows(.id = "Level") %>% 
  kable()
```

## 8-daily, spatial and annual

```{r message = FALSE, warning = FALSE, echo = FALSE}
out_eval$gpp$fluxnet$plot$gg_modobs_xdaily
out_eval$gpp$fluxnet$plot$gg_modobs_spatial_annual
```

## Mean seasonal cycle

```{r message = FALSE, warning = FALSE, echo = FALSE}
out_eval$gpp$fluxnet$data$meandoydf_byclim %>% 
  dplyr::filter(climatezone %in% c(
    "Aw south", "BSk north", "Cfa north",
    "Cfb north", "Cfb south", "Csa north",
    "Csb north", "Dfb north", "Dfc north")
    ) %>%
  dplyr::filter(koeppen_code != "-") %>% 
  pivot_longer(
    c(obs_mean, mod_mean),
    names_to = "source",
    values_to = "gpp"
    ) %>% 
  ggplot() +
  geom_ribbon(
    aes(x = doy, ymin = obs_min, ymax = obs_max), 
    fill = "black", 
    alpha = 0.2
    ) +
  geom_line(aes(x = doy, y = gpp, color = source), size = 0.4) +
  labs(y = expression( paste("Simulated GPP (g C m"^-2, " d"^-1, ")" ) ), 
       x = "DOY") +
  facet_wrap( ~climatezone ) +
  theme_gray() +
  theme(legend.position = "bottom") +
  scale_color_manual(
    name = "Setup: ",
    values = c("red", "black")
    )
```

## Drought response

```{r message = FALSE, warning = FALSE, echo = FALSE}
source("R/align_events.R")
source("R/eval_droughtresponse.R")

df_dday_agg <- eval_droughtresponse( 
  df = out_eval$gpp$fluxnet$data$ddf %>%
    rename(
      site = sitename
      ), 
  df_flue = readr::read_csv(here::here("data/flue_stocker18nphyt.csv")),
  before = 20,
  after = 105,
  leng_threshold = 10, 
  nbins = 10, 
  do_norm = TRUE
  )

df_dday_agg %>% 
  ggplot() +
  geom_hline(
    yintercept = 0,
    color = "black",
    linetype = "dotted"
    ) +
  geom_vline(
    xintercept = 0,
    color = "black",
    linetype = "dotted"
    ) +
  geom_line(
    aes(
      x = dday,
      y = median
      ),
    size = 0.9) +
  geom_ribbon(
    aes(
      x = dday,
      ymin = q33,
      ymax = q66
      ), 
    alpha = 0.3) +
  scale_color_manual(
    values = c(
      "BRC" = "black",
      "FULL" = "royalblue"
      ),
    name = "Setup"
    ) +
  scale_fill_manual(
    values = c(
      "BRC" = "black",
      "FULL" = "royalblue"
      ),
    name = "Setup") +
  ylim(-1.2, 2.2) + 
  xlim(-20, 105) +
  scale_x_continuous(
    expand = c(0,0)
    ) +
  scale_y_continuous(
    expand = c(0,0)
    ) +
  labs(
    x = "Days after drought onset",
    y = expression( paste( "Bias (g C m"^{-1}, " d"^{-1}, ")")) 
    ) +
  theme_classic() +
  theme(
    panel.grid.major.y = element_line(),
    panel.grid.major.x = element_line()
  )
```

## Appendix

```{r}
sessionInfo()
```
