

```{r setup, include = FALSE}
# Install packages
if ("rsofun" %in% installed.packages()) remove.packages("rsofun")
devtools::install_github(
  "geco-bern/rsofun",
  ref = "HEAD",
  upgrade = "never",
  force = TRUE
  )

packageVersion("rsofun")


# Load packages
library(rsofun)
library(dplyr)
library(tidyr)
library(readr)
library(ggplot2)
library(lubridate)
library(knitr)
library(ggthemes)
library(here)

# Load functions
source("../R/eval_sofun.R")
source("../R/get_stats.R")
source("../R/analyse_modobs2.R")
source(here("R/align_events.R"))
source(here("R/eval_droughtresponse.R"))
source(here("R/heatscatter_dependencies.R"))

# Set seed
set.seed(42)

# only select the sites from the New Phytologist paper
# load site list and filter on these (see below)
flue_sites <- readr::read_csv(here::here("data/flue_stocker18nphyt.csv")) %>%
  dplyr::filter( !is.na(cluster) ) %>%
  distinct(site) %>%
  pull(site)

# model forcing data
load(here("data/df_drivers_fluxnet2015.Rdata"))

df_drivers_fluxnet2015 <- df_drivers_fluxnet2015 %>%
  dplyr::select(sitename, forcing) %>%
  unnest(forcing) %>%
  dplyr::filter(!(month(date)==2 & mday(date)==29)) %>%

  ## model requires flux per seconds now
  mutate(prec = prec / (60*60*24), ppfd = ppfd / (60*60*24)) %>%

  ## assuming all precipitation in liquid form
  mutate(rainf = prec, snowf = 0) %>%

  ## required for new version, but not used because
  mutate(tmin = temp, tmax = temp) %>%

  group_by(sitename) %>%
  nest() %>%
  rename(forcing = data) %>%
  right_join(
    df_drivers_fluxnet2015 %>%
      dplyr::select(-forcing),
    by = "sitename"
  ) %>%
  ungroup() %>%
  rename(site_info = siteinfo, params_soil = df_soiltexture)

# site selection: good-quality sites from analysis of Stocker et al. (2018) New Phyt.
flue_sites <- readr::read_csv(here("data/flue_stocker18nphyt.csv")) %>%
  dplyr::filter( !is.na(cluster) ) %>%
  distinct(site) %>%
  pull(site)


ddf_fluxnet_gpp <- readr::read_rds(here("data/ddf_fluxnet_gpp_v4.2.rds"))

# observational data for evaluation
obs_eval <- readRDS(here("data","correct_2015_eval_data.rds"))

# site meta info
siteinfo_fluxnet2015 <- df_drivers_fluxnet2015 |>
  select(sitename, site_info) |>
  unnest(site_info)

# calibration sites
calibsites <- siteinfo_fluxnet2015 %>%
  dplyr::filter(!(sitename %in% c("DE-Akm", "IT-Ro1"))) %>%  # excluded because fapar data could not be downloaded (WEIRD)
  # dplyr::filter(!(sitename %in% c("AU-Wom"))) %>%  # excluded because no GPP data was found in FLUXNET file
  dplyr::filter(sitename != "FI-Sod") %>%  # excluded because some temperature data is missing
  dplyr::filter( c4 %in% c(FALSE, NA) & classid != "CRO" & classid != "WET" ) %>%
  dplyr::filter( sitename %in% flue_sites ) %>%
  pull(sitename)

# Calibrate model
siteinfo_fluxnet2015 <- df_drivers_fluxnet2015 |> 
  select(sitename, site_info) |> 
  unnest(site_info)

```

# Model parameters

Model parameters are determined for this version from calibration. The code for model calibration is not shown. See `vignettes/0_p-model_current_run.Rmd` for the code.

```{r}
par_calib <-  read_rds(here("data","calibration_old_data.rds"))

params_modl <- list(
    kphio              = par_calib$par[["kphio"]], 
    kphio_par_a        = par_calib$par[["kphio_par_a"]],       
    kphio_par_b        = par_calib$par[["kphio_par_b"]],
    soilm_thetastar    = par_calib$par[["soilm_thetastar"]],
    soilm_betao        = 0.0,
    beta_unitcostratio = 146.0,
    rd_to_vcmax        = 0.014,      # value from Atkin et al. 2015 for C3 herbaceous
    tau_acclim         = 30.0,
    kc_jmax            = 0.41
  )
```

# Model run

```{r warning = FALSE, message = FALSE}
output <- rsofun::runread_pmodel_f(
  df_drivers_fluxnet2015 |>
    dplyr::mutate(
      forcing = purrr::map(
        forcing,
        ~rename(., rain = rainf, snow = snowf))
      ) |>
    dplyr::mutate(
      forcing = purrr::map(
        forcing,
        ~mutate(., netrad = 0))
    ),
  par = params_modl
)

output <- output |>
  mutate(len = purrr::map_int(data, ~nrow(.))) |>
  filter(len != 1) |>
  select(-len)
```

# Run evaluation

This runs the `eval_sofun()` routine included in the sofunCalVal package, and outputs daily, monthly and annual summary statistics comparing observed with simulated values. We'll retake this same routine when running on the latest release (i.e. the workflow is equivalent but for the release of the `rsofun` package)

```{r warning=FALSE, message=FALSE, error=FALSE, include = FALSE}
evalsites <- output %>%
  mutate(ntsteps = purrr::map_dbl(data, ~nrow(.))) %>%
  dplyr::filter(ntsteps > 0) %>%
  pull(sitename)

settings_eval <- list(
  benchmark = list( gpp = c("fluxnet") ),
  sitenames = evalsites,
  agg       = 8
)

##  Evaluate model ----
out_eval <- eval_sofun(
  output,
  settings_eval,
  obs_eval = obs_eval,
  overwrite = TRUE,
  light = FALSE
)
```

# Results

```{r warning=FALSE, message=FALSE, error=FALSE, echo = FALSE}
out_eval$gpp$fluxnet$metrics %>% 
  bind_rows(.id = "Level") %>% 
  kable()
```

## 8-daily, spatial and annual

```{r message = FALSE, warning = FALSE, echo = FALSE}
out_eval$gpp$fluxnet$plot$gg_modobs_xdaily
out_eval$gpp$fluxnet$plot$gg_modobs_spatial_annual
```

## Mean seasonal cycle

```{r message = FALSE, warning = FALSE, echo = FALSE}
out_eval$gpp$fluxnet$data$meandoydf_byclim %>% 
  dplyr::filter(climatezone %in% c(
    "Aw south", "BSk north", "Cfa north",
    "Cfb north", "Cfb south", "Csa north",
    "Csb north", "Dfb north", "Dfc north")
    ) %>%
  dplyr::filter(koeppen_code != "-") %>% 
  pivot_longer(
    c(obs_mean, mod_mean),
    names_to = "source",
    values_to = "gpp"
    ) %>% 
  ggplot() +
  geom_ribbon(
    aes(x = doy, ymin = obs_min, ymax = obs_max), 
    fill = "black", 
    alpha = 0.2
    ) +
  geom_line(aes(x = doy, y = gpp, color = source), size = 0.4) +
  labs(y = expression( paste("Simulated GPP (g C m"^-2, " d"^-1, ")" ) ), 
       x = "DOY") +
  facet_wrap( ~climatezone ) +
  theme_gray() +
  theme(legend.position = "bottom") +
  scale_color_manual(
    name="Setup: ",
    values=c("red", "black")
    )
```

## Drought response

```{r message = FALSE, warning = FALSE, echo = FALSE}

df_dday_agg <- eval_droughtresponse( 
  df = out_eval$gpp$fluxnet$data$ddf %>%
    rename(
      site = sitename
      ), 
  df_flue = readr::read_csv(here::here("data/flue_stocker18nphyt.csv")),
  before=20,
  after=105,
  leng_threshold = 10, 
  nbins=10, 
  do_norm=TRUE
  )

df_dday_agg %>% 
  ggplot() +
  geom_hline(
    yintercept = 0,
    color = "black",
    linetype = "dotted"
    ) +
  geom_vline(
    xintercept = 0,
    color = "black",
    linetype = "dotted"
    ) +
  geom_line(
    aes(
      x = dday,
      y = median
      ),
    size = 0.9) +
  geom_ribbon(
    aes(
      x = dday,
      ymin = q33,
      ymax = q66
      ), 
    alpha = 0.3) +
  scale_color_manual(
    values = c(
      "BRC" = "black",
      "FULL" = "royalblue"
      ),
    name = "Setup"
    ) +
  scale_fill_manual(
    values = c(
      "BRC" = "black",
      "FULL" = "royalblue"
      ),
    name = "Setup") +
  ylim(-1.2, 2.2) + 
  xlim(-20, 105) +
  scale_x_continuous(
    expand = c(0,0)
    ) +
  scale_y_continuous(
    expand = c(0,0)
    ) +
  labs(
    x = "Days after drought onset",
    y = expression( paste( "Bias (g C m"^{-1}, " d"^{-1}, ")")) 
    ) +
  theme_classic()
```

```{r}
sessionInfo()
```
