```{r setup}

# Load packages
library(rsofun)
library(dplyr)
library(tidyr)
library(readr)
library(ggplot2)
library(lubridate)
library(knitr)
library(ggthemes)
library(here)

# Load functions
source(here("R/eval_sofun.R"))
source(here("R//get_stats.R"))
source(here("R//analyse_modobs2.R"))
source(here("R/align_events.R"))
source(here("R/eval_droughtresponse.R"))
source(here("R/heatscatter_dependencies.R"))
source(here("R/create_obs_eval.R"))


# Set seed
set.seed(42)

# model forcing data
driver <- read_rds("/data_2/FluxDataKit/v3.4/zenodo_upload/rsofun_driver_data_v3.4.2.rds") # currently only on workstation

# site meta info
fdk_site_info <- read_csv("/data_2/FluxDataKit/v3.4/zenodo_upload/fdk_site_info.csv")

# mispell
fdk_site_info$koeppen_code <-  ifelse(fdk_site_info$koeppen_code == "BSh","Bsh",fdk_site_info$koeppen_code)
fdk_site_info$koeppen_code <-  ifelse(fdk_site_info$koeppen_code == "BWh","Bwh",fdk_site_info$koeppen_code)
fdk_site_info$koeppen_code <-  ifelse(fdk_site_info$koeppen_code == "BSk","Bsk",fdk_site_info$koeppen_code)


# data quality filter info
fdk_filter <-  read_csv("/data_2/FluxDataKit/v3.4/zenodo_upload/fdk_site_fullyearsequence.csv")

# exclude croplands and wetlands from calibration/evaluation
fdk_site_info <- fdk_site_info[
  fdk_site_info$igbp_land_use != "CRO" 
  & fdk_site_info$igbp_land_use != "WET"
                               ,]
driver_forcing <- driver |> 
 dplyr::select(sitename, forcing) |> 
  unnest(cols = c(forcing))

driver <- driver_forcing |>
  left_join(
    fdk_filter |> 
     dplyr::select(
        sitename, 
        year_start = year_start_gpp, 
        year_end = year_end_gpp),
    by = join_by(sitename)
  ) |> 
  mutate(year = year(date)) |> 
  filter(year >= year_start & year <= year_end) |> 
 dplyr::select(-year_start, -year_end, -year) |> 
  group_by(sitename) |> 
  nest() |> 
  left_join(
    driver |> 
     dplyr::select(
        sitename, 
        site_info,
        params_siml
      ),
    by = join_by(sitename)
  ) |> 
  rename(forcing = data) |> 
 dplyr::select(sitename, params_siml, site_info, forcing) |> 
  ungroup()


# apply good year sequence data quality filter for LE
fdk_filter <- fdk_filter[fdk_filter$drop_le == "FALSE",]

driver <- driver[which(driver$sitename %in% fdk_site_info$sitename &
                       driver$sitename %in% fdk_filter$sitename),]

fdk_site_info <- fdk_site_info[which(fdk_site_info$sitename %in% driver$sitename),]

fdk_filter <- fdk_filter[which(fdk_filter$sitename %in% driver$sitename &
                               fdk_filter$sitename %in% fdk_site_info$sitename),]

driver_forcing <- driver |> 
 dplyr::select(sitename, forcing) |> 
  unnest(cols = c(forcing))

driver <- driver_forcing |>
  left_join(
    fdk_filter |> 
     dplyr::select(
        sitename, 
        year_start = year_start_le, 
        year_end = year_end_le),
    by = join_by(sitename)
  ) |> 
  mutate(year = year(date)) |> 
  filter(year >= year_start & year <= year_end) |> 
 dplyr::select(-year_start, -year_end, -year) |> 
  group_by(sitename) |> 
  nest() |> 
  left_join(
    driver |> 
     dplyr::select(
        sitename, 
        site_info,
        params_siml
      ),
    by = join_by(sitename)
  ) |> 
  rename(forcing = data) |> 
 dplyr::select(sitename, params_siml, site_info, forcing) |> 
  ungroup()

# sitedplyr::selection: good-quality sites from analysis of Stocker et al. (2018) New Phyt.
df_flue <- readr::read_csv("~/data_scratch/SoFunCalVal_definitivo/data/flue_stocker18nphyt.csv")

df_flue <- df_flue[df_flue$site %in% fdk_filter$sitename,]

df_flue <- left_join(df_flue |>
                       rename(sitename = site), fdk_filter |>
                      dplyr::select(sitename,year_start_gpp,year_end_gpp),
                       by = "sitename") |>
           filter(between(as_date(date),
                 as_date(paste0(year_start_gpp,"-01-01")),
                 as_date(paste0(year_end_gpp,"-12-31"))
                 )) |>
          dplyr::select(-c(year_start_gpp,year_end_gpp)) |>
           rename(site = sitename)

# filter out sites with ET higher than precipitation

le_to_et <- function(df){
  1000 * 60 * 60 * 24 * df$le / (cwd::calc_enthalpy_vap(df$temp) * cwd::calc_density_h2o(df$temp, df$patm))
}


filter <- driver |>
  unnest(forcing) |>
  dplyr::select(sitename,le,temp,patm,rain) |>
  mutate(rain = rain * 60 * 60 * 24 )

filter$et <- le_to_et(filter)

filter <- filter |>
  group_by(sitename) |>
  summarise(rain = sum(rain,na.rm = T),
            et = sum(et,na.rm = T)) |>
  mutate(to_drop = ifelse(et > rain,T,F))

filter <- filter |> filter(to_drop == F)

driver <- driver[driver$sitename %in% filter$sitename,]

fdk_site_info <- fdk_site_info[fdk_site_info$sitename %in% driver$sitename,]


driver <- driver |> 
  unnest(forcing) |>
  mutate(aet = 1000 * 60 * 60 * 24 * le / (cwd::calc_enthalpy_vap(temp) * cwd::calc_density_h2o(temp, patm))) |>
  nest(forcing = c("date","temp", "vpd", "ppfd", "netrad", "patm", "snow", "rain", "tmin", "tmax", "vwind", "fapar", "co2", "ccov",
                   "gpp", "gpp_qc", "nee", "nee_qc", "le", "le_qc", "aet"))

# construct list of observations used for evaluation
obs_eval <- create_obs_eval(driver ,fdk_site_info, target = c("gpp","aet"))
```


# Model run

```{r run, warning = FALSE, message = FALSE}
# best paramters

params_modl <- list(
  kphio              =  4.646905e-02, 
  kphio_par_a        = -9.999618e-04,       
  kphio_par_b        =  2.354870e+01,
  rd_to_vcmax        =  0.014,  
  soilm_thetastar    =  3.964174e+02, 
  beta_unitcostratio =  1.546904e+02,
  tau_acclim         =  30,
  kc_jmax            =  0.41,
  gw_calib           =  6.619571e-01
)

driver <- driver |> unnest(params_siml) |>
  mutate(use_gs = TRUE,
         use_phydro = FALSE,
         use_pml= TRUE) |>
  group_by(sitename) |>
  nest(params_siml = c(spinup, spinupyears, recycle, outdt, ltre,  ltne,  ltrd,  ltnd,  lgr3,  lgn3,  lgr4 ,
                       use_gs, use_phydro, use_pml))


 output <- rsofun::runread_pmodel_f(
  driver,
  par = params_modl
 )
```

# Run evaluation

This runs the `eval_sofun()` routine included in the sofunCalVal package, and outputs daily, monthly and annual summary statistics comparing observed with simulated values. We'll retake this same routine when running on the latest release (i.e. the workflow is equivalent but for the release of the `rsofun` package)

```{r evaluation, warning=FALSE, message=FALSE, error=FALSE}

settings_eval <- list(
  benchmark = list( gpp = c("fluxnet"), aet = c("fluxnet") ),
  sitenames = output$sitename,
  agg       = 7
  )

out_eval <- eval_sofun( 
  output, 
  settings_eval, 
  obs_eval = obs_eval, 
  overwrite = TRUE, 
  light = FALSE 
  )
```

# Results GPP

```{r warning=FALSE, message=FALSE, error=FALSE, echo = FALSE}
out_eval$gpp$fluxnet$metrics %>% 
  bind_rows(.id = "Level") %>% 
  kable(caption= "GPP")
```

## 8-daily, spatial and annual

```{r message = FALSE, warning = FALSE, echo = FALSE}
out_eval$gpp$fluxnet$plot$gg_modobs_xdaily
out_eval$gpp$fluxnet$plot$gg_modobs_spatial_annual
```

## Mean seasonal cycle

```{r message = FALSE, warning = FALSE, echo = FALSE}

climates <-  out_eval$gpp$fluxnet$data$meandoydf_byclim_stats

metrics_rsq <- climates |> 
  group_by(climatezone) |>
  summarise(rsq = mean(rsq), 
            rmse = mean(RMSE),# are repeated elements
            x_pos = 365,
            y_pos = 17.5) |>
  mutate(rsq = paste0("rsq = ", substr(rsq,1,4)))

metrics_rmse <- climates |> 
  group_by(climatezone) |>
  summarise( 
            rmse = mean(RMSE),# are repeated elements
            x_pos = 365,
            y_pos = 14.5) |>
  mutate(rmse = paste0("rmse = ", substr(rmse,1,4)))


out_eval$gpp$fluxnet$data$meandoydf_byclim %>% 
  dplyr::filter(climatezone %in%  unique(out_eval$gpp$fluxnet$data$meandoydf_byclim$climatezone)[1:10]
    ) %>%
  dplyr::filter(climatezone != "- north") %>% 
  pivot_longer(
    c(obs_mean, mod_mean),
    names_to = "source",
    values_to = "gpp"
    ) %>% 
  ggplot() +
  geom_ribbon(
    aes(x = doy, ymin = obs_min, ymax = obs_max), 
    fill = "black", 
    alpha = 0.2
    ) +
  geom_line(aes(x = doy, y = gpp, color = source), size = 0.4) +
  labs(y = expression( paste("Simulated GPP (g C m"^-2, " d"^-1, ")" ) ), 
       x = "DOY") +
  facet_wrap( ~climatezone ) +
  geom_text(data = metrics_rsq |> dplyr::filter(climatezone %in%  unique(out_eval$gpp$fluxnet$data$meandoydf_byclim$climatezone)[2:10]), 
                                                aes(x = x_pos, y = y_pos, label = rsq), size = 4, hjust = 1) +
  geom_text(data = metrics_rmse |> dplyr::filter(climatezone %in%  unique(out_eval$gpp$fluxnet$data$meandoydf_byclim$climatezone)[2:10]), 
                                               aes(x = x_pos, y = y_pos, label = rmse), size = 4, hjust = 1) +
  theme_gray() +
  theme(legend.position = "bottom") +
  scale_color_manual(
    name = "Setup: ",
    values = c("red", "black")
    )
```

```{r}
out_eval$gpp$fluxnet$data$meandoydf_byclim %>% 
  dplyr::filter(climatezone %in%  unique(out_eval$gpp$fluxnet$data$meandoydf_byclim$climatezone)[11:19]) %>%
  dplyr::filter(koeppen_code != "-") %>% 
  pivot_longer(
    c(obs_mean, mod_mean),
    names_to = "source",
    values_to = "gpp"
    ) %>% 
  ggplot() +
  geom_ribbon(
    aes(x = doy, ymin = obs_min, ymax = obs_max), 
    fill = "black", 
    alpha = 0.2
    ) +
  geom_line(aes(x = doy, y = gpp, color = source), size = 0.4) +
  labs(y = expression( paste("Simulated GPP (g C m"^-2, " d"^-1, ")" ) ), 
       x = "DOY") +
  facet_wrap( ~climatezone ) +
  geom_text(data = metrics_rsq |> dplyr::filter(climatezone %in%  unique(out_eval$gpp$fluxnet$data$meandoydf_byclim$climatezone)[11:19]), 
                                                aes(x = x_pos, y = y_pos, label = rsq), size = 4.5, hjust = 1) +
  geom_text(data = metrics_rmse |> dplyr::filter(climatezone %in%  unique(out_eval$gpp$fluxnet$data$meandoydf_byclim$climatezone)[11:19]), 
                                               aes(x = x_pos, y = y_pos, label = rmse), size = 4.5, hjust = 1) +
  theme_gray() +
  theme(legend.position = "bottom") +
  scale_color_manual(
    name = "Setup: ",
    values = c("red", "black")
    )
```

```{r}
out_eval$gpp$fluxnet$data$meandoydf_byclim %>%
  dplyr::filter(climatezone %in%  unique(out_eval$gpp$fluxnet$data$meandoydf_byclim$climatezone)[20:28]) %>%
  dplyr::filter(koeppen_code != "-") %>%
  pivot_longer(
    c(obs_mean, mod_mean),
    names_to = "source",
    values_to = "gpp"
    ) %>%
  ggplot() +
  geom_ribbon(
    aes(x = doy, ymin = obs_min, ymax = obs_max),
    fill = "black",
    alpha = 0.2
    ) +
  geom_line(aes(x = doy, y = gpp, color = source), size = 0.4) +
  labs(y = expression( paste("Simulated GPP (g C m"^-2, " d"^-1, ")" ) ),
       x = "DOY") +
  facet_wrap( ~climatezone ) +
    geom_text(data = metrics_rsq |> dplyr::filter(climatezone %in%  unique(out_eval$gpp$fluxnet$data$meandoydf_byclim$climatezone)[20:28]), 
                                                aes(x = x_pos, y = y_pos, label = rsq), size = 4.5, hjust = 1) +
  geom_text(data = metrics_rmse |> dplyr::filter(climatezone %in%  unique(out_eval$gpp$fluxnet$data$meandoydf_byclim$climatezone)[20:28]), 
                                               aes(x = x_pos, y = y_pos, label = rmse), size = 4.5, hjust = 1) +
  theme_gray() +
  theme(legend.position = "bottom") +
  scale_color_manual(
    name = "Setup: ",
    values = c("red", "black")
    )
```


## Drought response

```{r message = FALSE, warning = FALSE, echo = FALSE}
df_dday_agg <- eval_droughtresponse( 
  df = out_eval$gpp$fluxnet$data$ddf %>%
    rename(
      site = sitename
      ), 
  df_flue = readr::read_csv("~/data_scratch/SoFunCalVal_definitivo/data/flue_stocker18nphyt.csv"),
  before = 20,
  after = 105,
  leng_threshold = 10, 
  nbins = 10, 
  do_norm = TRUE
  )

df_dday_agg %>% 
  ggplot() +
  geom_hline(
    yintercept = 0,
    color = "black",
    linetype = "dotted"
    ) +
  geom_vline(
    xintercept = 0,
    color = "black",
    linetype = "dotted"
    ) +
  geom_line(
    aes(
      x = dday,
      y = median
      ),
    size = 0.9) +
  geom_ribbon(
    aes(
      x = dday,
      ymin = q33,
      ymax = q66
      ), 
    alpha = 0.3) +
  scale_color_manual(
    values = c(
      "BRC" = "black",
      "FULL" = "royalblue"
      ),
    name = "Setup"
    ) +
  scale_fill_manual(
    values = c(
      "BRC" = "black",
      "FULL" = "royalblue"
      ),
    name = "Setup") +
  ylim(-1.2, 2.2) + 
  xlim(-20, 105) +
  scale_x_continuous(
    expand = c(0,0)
    ) +
  scale_y_continuous(
    expand = c(0,0)
    ) +
  labs(
    x = "Days after drought onset",
    y = expression( paste( "Bias (g C m"^{-1}, " d"^{-1}, ")")) 
    ) +
  theme_classic() +
  theme(
    panel.grid.major.y = element_line(),
    panel.grid.major.x = element_line()
  )
```



# Results AET

```{r warning=FALSE, message=FALSE, error=FALSE, echo = FALSE}
out_eval$aet$fluxnet$metrics %>% 
  bind_rows(.id = "Level") %>% 
  kable(caption= "AET")
```

## 8-daily, spatial and annual

<!-- xxx plot means across weeks, months, years, ... -->

```{r message = FALSE, warning = FALSE, echo = FALSE}
out_eval$aet$fluxnet$plot$gg_modobs_xdaily
out_eval$aet$fluxnet$plot$gg_modobs_spatial_annual
```

## Mean seasonal cycle


```{r message = FALSE, warning = FALSE, echo = FALSE}

climates <-  out_eval$aet$fluxnet$data$meandoydf_byclim_stats

metrics_rsq <- climates |> 
  group_by(climatezone) |>
  summarise(rsq = mean(rsq), 
            rmse = mean(RMSE),# are repeated elements
            x_pos = 365,
            y_pos = 8.1) |>
  mutate(rsq = paste0("rsq = ", substr(rsq,1,4)))

metrics_rmse <- climates |> 
  group_by(climatezone) |>
  summarise( 
            rmse = mean(RMSE),# are repeated elements
            x_pos = 365,
            y_pos = 6.8) |>
  mutate(rmse = paste0("rmse = ", substr(rmse,1,4)))


out_eval$aet$fluxnet$data$meandoydf_byclim %>% 
  dplyr::filter(climatezone %in%  unique(out_eval$aet$fluxnet$data$meandoydf_byclim$climatezone)[2:10]) %>%
  dplyr::filter(koeppen_code != "-") %>% 
  pivot_longer(
    c(obs_mean, mod_mean),
    names_to = "source",
    values_to = "aet"
    )%>% 
  ggplot() +
  geom_ribbon(
    aes(x = doy, ymin = obs_min, ymax = obs_max), 
    fill = "black", 
    alpha = 0.2
    ) +
  geom_line(aes(x = doy, y = aet, color = source), size = 0.4) +
  labs(y =  expression( paste("Simulated AET (mm d"^{-1}, ")" ) ), 
       x = "DOY") +
  ylim(0,8.5) +
  facet_wrap( ~climatezone ) +
  geom_text(data = metrics_rsq |> dplyr::filter(climatezone %in%  unique(out_eval$aet$fluxnet$data$meandoydf_byclim$climatezone)[2:10]), 
                                                aes(x = x_pos, y = y_pos, label = rsq), size = 4.5, hjust = 1) +
  geom_text(data = metrics_rmse |> dplyr::filter(climatezone %in%  unique(out_eval$aet$fluxnet$data$meandoydf_byclim$climatezone)[2:10]), 
                                               aes(x = x_pos, y = y_pos, label = rmse), size = 4.5, hjust = 1) +
  theme_gray() +
  theme(legend.position = "bottom") +
  scale_color_manual(
    name = "Setup: ",
    values = c("red", "black")
    )


#ggsave("../prova.png", plot = plt, height = 1768, width = 2500 , units = "px",scale = 1)
```



```{r}
out_eval$aet$fluxnet$data$meandoydf_byclim %>% 
  dplyr::filter(climatezone %in%  unique(out_eval$aet$fluxnet$data$meandoydf_byclim$climatezone)[11:19]) %>%
  dplyr::filter(koeppen_code != "-") %>% 
  pivot_longer(
    c(obs_mean, mod_mean),
    names_to = "source",
    values_to = "aet"
    )%>% 
  ggplot() +
  geom_ribbon(
    aes(x = doy, ymin = obs_min, ymax = obs_max), 
    fill = "black", 
    alpha = 0.2
    ) +
  geom_line(aes(x = doy, y = aet, color = source), size = 0.4) +
  labs(y =  expression( paste("Simulated AET (mm d"^{-1}, ")" ) ), 
       x = "DOY") +
  ylim(0,8.5) +
  facet_wrap( ~climatezone ) +
  geom_text(data = metrics_rsq |> dplyr::filter(climatezone %in%  unique(out_eval$aet$fluxnet$data$meandoydf_byclim$climatezone)[11:19]), 
                                                aes(x = x_pos, y = y_pos, label = rsq), size = 4.5, hjust = 1) +
  geom_text(data = metrics_rmse |> dplyr::filter(climatezone %in%  unique(out_eval$aet$fluxnet$data$meandoydf_byclim$climatezone)[11:19]), 
                                               aes(x = x_pos, y = y_pos, label = rmse), size = 4.5, hjust = 1) +
  theme_gray() +
  theme(legend.position = "bottom") +
  scale_color_manual(
    name = "Setup: ",
    values = c("red", "black")
    )
```

```{r}
out_eval$aet$fluxnet$data$meandoydf_byclim %>% 
  dplyr::filter(climatezone %in%  unique(out_eval$aet$fluxnet$data$meandoydf_byclim$climatezone)[20:28]) %>%
  dplyr::filter(koeppen_code != "-") %>% 
  pivot_longer(
    c(obs_mean, mod_mean),
    names_to = "source",
    values_to = "aet"
    )%>% 
  ggplot() +
  geom_ribbon(
    aes(x = doy, ymin = obs_min, ymax = obs_max), 
    fill = "black", 
    alpha = 0.2
    ) +
  geom_line(aes(x = doy, y = aet, color = source), size = 0.4) +
  labs(y =  expression( paste("Simulated AET (mm d"^{-1}, ")" ) ), 
       x = "DOY") +
  ylim(0,8.5) +
  facet_wrap( ~climatezone ) +
  geom_text(data = metrics_rsq |> dplyr::filter(climatezone %in%  unique(out_eval$aet$fluxnet$data$meandoydf_byclim$climatezone)[20:28]), 
                                                aes(x = x_pos, y = y_pos, label = rsq), size = 4.5, hjust = 1) +
  geom_text(data = metrics_rmse |> dplyr::filter(climatezone %in%  unique(out_eval$aet$fluxnet$data$meandoydf_byclim$climatezone)[20:28]), 
                                               aes(x = x_pos, y = y_pos, label = rmse), size = 4.5, hjust = 1) +
  theme_gray() +
  theme(legend.position = "bottom") +
  scale_color_manual(
    name = "Setup: ",
    values = c("red", "black")
    )
```


## Drought response

```{r message = FALSE, warning = FALSE, echo = FALSE}
df_dday_agg <- eval_droughtresponse( 
  df = out_eval$aet$fluxnet$data$ddf %>%
    rename(
      site = sitename
      ), 
  df_flue = readr::read_csv("~/data_scratch/SoFunCalVal_definitivo/data/flue_stocker18nphyt.csv"),
  before = 20,
  after = 105,
  leng_threshold = 10, 
  nbins = 10, 
  do_norm = TRUE
  )

df_dday_agg %>% 
  ggplot() +
  geom_hline(
    yintercept = 0,
    color = "black",
    linetype = "dotted"
    ) +
  geom_vline(
    xintercept = 0,
    color = "black",
    linetype = "dotted"
    ) +
  geom_line(
    aes(
      x = dday,
      y = median
      ),
    size = 0.9) +
  geom_ribbon(
    aes(
      x = dday,
      ymin = q33,
      ymax = q66
      ), 
    alpha = 0.3) +
  scale_color_manual(
    values = c(
      "BRC" = "black",
      "FULL" = "royalblue"
      ),
    name = "Setup"
    ) +
  scale_fill_manual(
    values = c(
      "BRC" = "black",
      "FULL" = "royalblue"
      ),
    name = "Setup") +
  ylim(-1.2, 2.2) + 
  xlim(-20, 105) +
  scale_x_continuous(
    expand = c(0,0)
    ) +
  scale_y_continuous(
    expand = c(0,0)
    ) +
  labs(
    x = "Days after drought onset",
    y = expression( paste( "Bias AET (mm d"^{-1}, ")")) 
    ) +
  theme_classic() +
  theme(
    panel.grid.major.y = element_line(),
    panel.grid.major.x = element_line()
  )
```



